data_collation

### 2024-03-21 scripts/data_collation.Rmd
commbined data frames by category, some questions for cleaning up dataa for austin

# Load libraries 
```{r}
library(dplyr)
library(tidyverse)
```

@@

# Define categories
categories <- c("UPC", "swath", "Morph", "Site", "QUAD", "kelp", "fish", "COVER")

# Initialize an empty list to store data frames for each category
category_data_frames <- lapply(categories, function(cat) list())

# Read files into R and store as data frames
read_files <- function(files, category) {
  data_frames <- list()
  for (file in files) {
    # Read each file
    filepath <- file.path(directory, file)
    # Assuming your files are CSV files, adjust accordingly if they're different
    data <- tryCatch({
      read.csv(filepath, stringsAsFactors = FALSE, fileEncoding = "UTF-8")
    }, warning = function(w) {
      # Print the warning message
      print(paste("Warning while reading file:", filepath))
      print(w)
      # Return NULL if there's a warning
      return(NULL)
    }, error = function(e) {
      # Print the error message
      print(paste("Error reading file:", filepath))
      print(e)
      # Return NULL if there's an error
      return(NULL)
    })

    if (!is.null(data)) {
      # Extract subject from the file name
      subject <- gsub(".csv", "", file)  # Remove the file extension
      subject <- gsub(paste0(category, "_"), "", subject, fixed = TRUE)  # Remove category prefix

      # Store each data frame with the subject name as the key
      if (!subject %in% names(data_frames)) {
        data_frames[[subject]] <- data
      } else {
        # If subject already exists, combine the data frames
        data_frames[[subject]] <- rbind(data_frames[[subject]], data)
      }
    }
  }
  return(data_frames)
}

# Read files for each category into R and store as data frames
for (category in categories) {
  category_files <- files[grep(category, files)]
  category_data_frames[[category]] <- read_files(category_files, category)
}

# Now you have a list of lists of data frames for each category, combined by subjects in the title

# Create individual data frames for each category
UPC_df <- combined_data_frames[["UPC"]]
swath_df <- combined_data_frames[["swath"]]
Morph_df <- combined_data_frames[["Morph"]]
Site_df <- combined_data_frames[["Site"]]
QUAD_df <- combined_data_frames[["QUAD"]]
kelp_df <- combined_data_frames[["kelp"]]
fish_df <- combined_data_frames[["fish"]]
COVER_df <- combined_data_frames[["COVER"]]

@@

>>>>

# Load libraries 
```{r}
library(plyr)
library(dplyr)
library(tidyverse)
```

@@

# Initialize an empty list to store data frames
data_frames <- list()

# Loop through each file
for (file in files) {
    # Read the file into a data frame
    # Adjust the read function according to your file type (e.g., read.csv, read.table)
    # Specify other parameters as needed (e.g., header = TRUE)
    file_path <- file.path(directory, file)
    df_name <- gsub("\\..*", "", file)  # Extract file name without extension
    assign(df_name, read.csv(file_path, stringsAsFactors = FALSE))
}

# Now, each file is a separate data frame with its own variable name

# COVER data frame is its own category, is this supposed to combine with another? it was labeled, Kelp_COVER

@@

# Combine fish dataframes
```{r}

# use rbind.fill() method in R to combine data frames with different columns
fish_data <- rbind.fill(FtWeatherill_fish_2016, FtWeatherill_fish_2017, fish_2018_Hump_Lab_size_classes, fish, fish_Aggregated)

# Combine UPC dataframes
```{r}

# use rbind.fill() method in R to combine data frames with different columns
UPC_data <- rbind.fill(UPC, UPC_2017, UPC_2018, FtWeatherill_UPC_2016, FtWeatherill_UPC_2016_CT)

# Combine swath dataframes
```{r}

# use rbind.fill() method in R to combine data frames with different columns
swath_data <- rbind.fill(swath, swath_2018, swath_count, FtWeatherill_swath_2016)

# the one swath data.frame has 115 variables because each species is listed instead of its code, fo you want me to replace with the code and make it wide to long format?

# Combine Kelp_Morph dataframes
```{r}

# use rbind.fill() method in R to combine data frames with different columns
morph_data <- rbind.fill(Subsurface_Kelp_Morph_2017, Subsurface_Kelp_Morph_2018)

# Combine site.info dataframes
```{r}

# use rbind.fill() method in R to combine data frames with different columns
site.info.dat <- rbind.fill(Site_info_2017, Site_info_2018, Site_info_template_v1)

# Combine site.info dataframes
```{r}

# use rbind.fill() method in R to combine data frames with different columns
site.info.dat <- rbind.fill(Site_info_2017, Site_info_2018, Site_info_template_v1)

# Combine QUAD dataframes
```{r}

# use rbind.fill() method in R to combine data frames with different columns
QUAD_data <- rbind.fill(QUAD, QUAD_2018, FtWeatherill_QUAD_2017, FtWeatherill_QUAD_2016_TALBENHORIN)

# the one QUAD data.frame has 130 variables because each species is listed instead of its code, fo you want me to replace with the code and make it wide to long format?

# Combine kelp dataframes
```{r}

# use rbind.fill() method in R to combine data frames with different columns
kelp_data <- rbind.fill(kelp, FtWeatherill_kelp_2016)

#delete repetative data collector columns
kelp_data <- kelp_data[ -c(13:16) ]
